Feature Selection - Boruta

### 2020-04-09 poc/EDA/EDA.Rmd

library(kableExtra)      # help you build common complex tables and manipulate table styles
library(tidyverse)       # for general data wrangling (includes readr and dplyr)
library(ggplot2)         # to draw statistical plots 
library(plotly)          # to construct interactive 3d plots
library(DataExplorer)    # automated data exploration
library(corrplot)        # to plot nice correlation matrix
library(caret)           # includes several functions to pre-process
library(scales)          # to determining breaks and labels for axes and legends
library(skimr)
library(funModeling) 
library(Hmisc)
library(grid)
library(hrbrthemes)
library(tidyr)
library(viridis)
library(ggpubr)
library(ggthemes)
library(GGally)
library(nortest)
library(emdist)
```

@@

#### Training

## QQ plot in R to compare two data samples
for (i in user_eng) {
  x <- df_beta_ue[,i]
  y <- df_release_ue[,i]

  rg <- range(x, y, na.rm=T)

  test <- ks.test(x, y)$statistic
  # pvalue <- ks.test(x, y)$p.value

  test <- paste("KS Test = ", round(test, 3))
  # pvalue <- paste("P-value = ", round(pvalue, 3))

  qqplot(x, y, main=i, xlim=rg, ylim=rg, xlab = "Beta", ylab = "Release", pch = 1)
  # mtext(test, side=3)
  text(min(x), max(x), test, adj=c(0,1))
  abline(0,1, col="red", lty=2)  
}
```

#### Validation

## QQ plot in R to compare two data samples
for (i in user_eng) {
  x <- df_beta_v_ue[,i]
  y <- df_release_v_ue[,i]

  rg <- range(x, y, na.rm=T)

  test <- ks.test(x, y)$statistic
  # pvalue <- ks.test(x, y)$p.value

  test <- paste("KS Test = ", round(test, 3))
  # pvalue <- paste("P-value = ", round(pvalue, 3))

  qqplot(x, y, main=i, xlim=rg, ylim=rg, xlab = "Beta", ylab = "Release", pch = 1)
  # mtext(test, side=3)
  text(min(x), max(x), test, adj=c(0,1))
  abline(0,1, col="red", lty=2)  
}
```

@@

@@

  rel_beta <- table(x)/nrow(df_beta_ue_dis) #divide the frequency counts by the total
  beta_bar <- barplot(rel_beta,
        main = "Beta", #Give your chart a title
        ylim=c(0,1), border=F, col = "#ffd271",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  ) 
  # Add the text 
  text(beta_bar, rel_beta+0.025, paste(round(rel_beta*100), "%", sep="") ,cex=1)

  rel_release <- table(y)/nrow(df_release_ue_dis) #divide the frequency counts by the total
  release_bar <- barplot(rel_release,
        main = "Release", #Give your chart a title
        ylim=c(0,1), border=F, col = "#f65c78",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  )

@@

  rel_beta <- table(x)/nrow(df_beta_v_ue_dis) #divide the frequency counts by the total
  beta_bar <- barplot(rel_beta,
        main = "Beta", #Give your chart a title
        ylim=c(0,1), border=F, col = "#ffd271",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  ) 
  # Add the text 
  text(beta_bar, rel_beta+0.025, paste(round(rel_beta*100), "%", sep="") ,cex=1)

  rel_release <- table(y)/nrow(df_release_v_ue_dis) #divide the frequency counts by the total
  release_bar <- barplot(rel_release,
        main = "Release", #Give your chart a title
        ylim=c(0,1), border=F, col = "#f65c78",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  )

>>>>

library(kableExtra)      # help you build common complex tables and manipulate table styles
library(tidyverse)       # for general data wrangling (includes readr and dplyr)
library(ggplot2)         # to draw statistical plots 
library(plotly)          # to construct interactive 3d plots
library(DataExplorer)    # automated data exploration
library(corrplot)        # to plot nice correlation matrix
library(caret)           # includes several functions to pre-process
library(scales)          # to determining breaks and labels for axes and legends
library(skimr)
library(funModeling) 
library(Hmisc)
library(grid)
library(hrbrthemes)
library(tidyr)
library(viridis)
library(ggpubr)
library(ggthemes)
library(GGally)
library(nortest)
library(emdist)
library(ggridges)
library(cowplot)
```

@@

```{r fig.height=9, fig.width=9, message=FALSE, warning=FALSE}
par(mfrow = c(4, 2))  ## Set up a 2 x 3 plotting space

## QQ plot in R to compare two data samples
for (i in user_eng) {

  # Training
  x_t <- df_beta_ue[,i]
  y_t <- df_release_ue[,i]

  rg_t <- range(x_t, y_t, na.rm=T)

  test_t <- ks.test(x_t, y_t)$statistic
  test_t <- paste("KS Test = ", round(test_t, 3))

  # Validation

  x_v <- df_beta_v_ue[,i]
  y_v <- df_release_v_ue[,i]

  rg_v <- range(x_v, y_v, na.rm=T)

  test_v <- ks.test(x_v, y_v)$statistic
  test_v <- paste("KS Test = ", round(test_v, 3))

  ########

  title_t <- paste('V67', i, sep='\n')
  qqplot(x_t, y_t, main=title_t, xlim=rg_t, ylim=rg_t, xlab = "Beta", ylab = "Release", pch = 1)
  # mtext(test, side=3)
  text(min(x_t), (if(max(x_t) > max(y_t)) max(x_t) else max(y_t)), test_t, adj=c(0,1))
  abline(0,1, col="#fe346e", lty=2)

  title_v <- paste('V68', i, sep='\n')
  qqplot(x_v, y_v, main=title_v, xlim=rg_v, ylim=rg_v, xlab = "Beta", ylab = "Release", pch = 1)
  # mtext(test, side=3)
  text(min(x_v), (if(max(x_v) > max(y_v)) max(x_v) else max(y_v)), test_v, adj=c(0,1))
  abline(0,1, col="#fe346e", lty=2)  
}
```

@@

### Violin Plots

The following violin plots depicts distributions for the beta and release subsets, for both versions v67 and v68. Violin plot is a powerful data visualization technique since it allows to compare both the **ranking** of several groups and their **distribution**.

**NOTE:** Guiding lines have been added for the following:

* <span style="color:red">red</span> dashed: Release mean
* <span style="color:blue">blue</span> dashed: Beta mean
* <span style="color:red">red</span> solid: Release median
* <span style="color:blue">blue</span> solid: Beta median

```{r fig.height=10, fig.width=15, message=FALSE, warning=FALSE}

## Violin plots
for (i in user_eng) {
  df_validate_means <- df_validate_ue %>%
  group_by(label) %>%
  summarise(value = mean(eval(as.name(i))))

  df_validate_medians <- df_validate_ue %>%
  group_by(label) %>%
  summarise(value = median(eval(as.name(i))))

  df_train_means <- df_train_ue %>%
  group_by(label) %>%
  summarise(value = mean(eval(as.name(i))))

  df_train_medians <- df_train_ue %>%
  group_by(label) %>%
  summarise(value = median(eval(as.name(i))))

  ###########

  violin_train <- ggplot(df_train_ue, aes(x=label, y=eval(as.name(i)), fill=label)) + 
  geom_violin(trim=FALSE) +
  labs(title=i,x="Channel", y = "Measure") + 
  scale_fill_manual(values=c("#111d5e", "#b21f66")) +
  geom_hline(
    data = df_train_means,
    aes(yintercept = value, colour = label),
    linetype = "dashed",
    size = 1
  ) +
  geom_hline(
    data = df_train_medians,
    aes(yintercept = value, colour = label),
    linetype = "solid",
    size = 1
  ) + scale_colour_manual(values=c("blue", "red")) +
  geom_boxplot(width=0.1, fill="white", alpha=0.5) + 
  theme_minimal() + theme(legend.position="none") + coord_flip()

  violin_valid <- ggplot(df_validate_ue, aes(x=label, y=eval(as.name(i)), fill=label)) + 
  geom_violin(trim=FALSE) +
  labs(title=i,x="Channel", y = "Measure") + 
  scale_fill_manual(values=c("#111d5e", "#b21f66")) +
  geom_hline(
    data = df_validate_means,
    aes(yintercept = value, colour = label),
    linetype = "dashed",
    size = 1
  ) +
  geom_hline(
    data = df_validate_medians,
    aes(yintercept = value, colour = label),
    linetype = "solid",
    size = 1
  ) + scale_colour_manual(values=c("blue", "red")) +
  geom_boxplot(width=0.1, fill="white", alpha=0.5) + 
  theme_minimal() + theme(legend.position="none") + coord_flip()

  print(plot_grid(violin_train, violin_valid, ncol = 1,labels = c('V67','V68')))
}
```

The violin plots above show the relationship of channel type to user engagement metrics. Overall, the results were quite similar comparing the two versions (v67 and v68). Only the following metrics yielded different results:

* `active_hours_max`
* `daily_unique_domains_max`

Now, let's take a closer look at the similar comparative results:

> Which variables presented **similar/equal** means and medians for beta and release users?

* `uri_count`
* `uri_count_max`
* `session_length_max`
* `search_count`
* `search_count_max`
* `daily_unique_domains`
* `daily_tabs_opened_max`
* `install_year`
* `profile_age`
* `memory_mb`
* `cpu_speed_mhz`

> Which variables presented **higher** release users means and/or medians than betas?

* `num_active_days`
* `active_hours`
* `num_pages`
* `num_pages_max`
* `daily_num_sessions_started`
* `daily_num_sessions_started_max`
* `cpu_cores`

> Which variables presented **lower** release users means and/or medians than betas?

* `session_length`
* `num_addons`
* `daily_tabs_opened`
* `timezone_offset`
* `cpu_l2_cache_kb`

Four variables became difficult to visualize through the violin plot. So we will plot ridgeline charts.

```{r fig.height=10, fig.width=15, message=FALSE, warning=FALSE}
vlist <- c('num_bookmarks','daily_max_tabs','daily_max_tabs_max','startup_ms')

## Ridgeline  plots
for (i in vlist) {
  ridge_train <- ggplot(df_train_ue, aes(x = eval(as.name(i)), y = label, fill = label)) +
  geom_density_ridges(alpha=0.6, stat="binline", bins=20) +
  labs(title=i,x="Channel", y = "Measure") + 
  theme_ridges() + 
  scale_fill_manual(values=c("#111d5e", "#b21f66")) +
  theme(legend.position = "none")

  ridge_valid <- ggplot(df_validate_ue, aes(x = eval(as.name(i)), y = label, fill = label)) +
  geom_density_ridges(alpha=0.6, stat="binline", bins=20) +
  labs(title=i,x="Channel", y = "Measure") + 
  theme_ridges() + 
  scale_fill_manual(values=c("#111d5e", "#b21f66")) +
  theme(legend.position = "none")

  print(plot_grid(ridge_train, ridge_valid, ncol = 1,labels = c('V67','V68')))
}
```

@@

  rel_beta <- table(x)/nrow(df_beta_ue_dis) #divide the frequency counts by the total
  beta_bar <- barplot(rel_beta,
        main = "Beta", #Give your chart a title
        ylim=c(0,1), border=F, col = "#111d5e",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  ) 
  # Add the text 
  text(beta_bar, rel_beta+0.025, paste(round(rel_beta*100), "%", sep="") ,cex=1)

  rel_release <- table(y)/nrow(df_release_ue_dis) #divide the frequency counts by the total
  release_bar <- barplot(rel_release,
        main = "Release", #Give your chart a title
        ylim=c(0,1), border=F, col = "#b21f66",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  )

@@

  rel_beta <- table(x)/nrow(df_beta_v_ue_dis) #divide the frequency counts by the total
  beta_bar <- barplot(rel_beta,
        main = "Beta", #Give your chart a title
        ylim=c(0,1), border=F, col = "#111d5e",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  ) 
  # Add the text 
  text(beta_bar, rel_beta+0.025, paste(round(rel_beta*100), "%", sep="") ,cex=1)

  rel_release <- table(y)/nrow(df_release_v_ue_dis) #divide the frequency counts by the total
  release_bar <- barplot(rel_release,
        main = "Release", #Give your chart a title
        ylim=c(0,1), border=F, col = "#b21f66",
        xlab = i, #Label the x axis
        ylab = "Relative Frequency" #Label the y axis
  )

### 2020-04-09 poc/matchIt/feature_selection.Rmd

```{r data_load, echo=FALSE}
load("~/GitHub/ff-beta-release-matching/poc/matchIt/feature_engineering.RData")
```

@@

set.seed(1234)
df_train_sm <- df_train_encoder %>%
   sample_n(1000)

df_fs <- df_train_sm %>%
   select(-engagement) %>%
   select(-content_crashes) %>%
   select(-client_id) %>%
   select(-label_beta) %>%
   select(-label_release) %>%
   select(-is_release) %>%
   select(-app_version)
```

@@

```{r apply_boruta}
boruta_results <- list()
for (metric in engagement){
  print(paste('Applying Boruta to ', metric))
  boruta.out <- Boruta(y = df_train_sm[[metric]], x=df_fs, doTrace=0)
  boruta_results[[metric]]  <- boruta.out
  # plot(boruta.out, cex.axis=.7, las=2, xlab="", main=metric) 
}
```

```{r boruta_plots}
for (metric in names(boruta_results)){
  plot(boruta_results[[metric]], cex.axis=.7, las=2, xlab="", main=metric) 
}
```

Find the top 5 ranking features per metric, and add to a list.

```{r aggregate_features_5}
features_top5 <- NULL
for(metric in names(boruta_results)){
  features_top5 <- c(names(sort(apply(boruta_results[[metric]]$ImpHistory, 2, median), decreasing = TRUE)[1:5]), features_top5)
}
knitr::kable(data.frame(top5 = sort(c(unique(features_top5)))))
```

Increasing to 10:

```{r aggregate_features_10}
features_top10 <- NULL
for(metric in names(boruta_results)){
  features_top10 <- c(names(sort(apply(boruta_results[[metric]]$ImpHistory, 2, median), decreasing = TRUE)[1:10]), features_top10)
}
knitr::kable(data.frame(top10 = sort(c(unique(features_top10)))))
```

@@

```{r var_def_equal}
df_beta <- df_train_encoder %>% 
  filter(label_beta == 1)
n_beta <- nrow(df_beta)
set.seed(1234)
df_rel <- df_train_encoder %>% 
  filter(label_beta == 0) %>%
  sample_n(n_beta)
set.seed(1234)
df_train_f_sm_eq <- df_rel %>%
  bind_rows(df_beta) %>%
   sample_n(1000)
df_fs_eq <- df_train_f_sm_eq %>%
   select(-engagement) %>%
   select(-content_crashes) %>%
   select(-client_id) %>%
   select(-label_beta) %>%
   select(-label_release) %>%
   select(-is_release) %>%
   select(-app_version)
```

```{r apply_boruta_equal}
boruta_results_eq <- list()
for (metric in engagement){
  print(paste('Applying Boruta to ', metric))
  boruta.out <- Boruta(y = df_train_f_sm_eq[[metric]], x=df_fs_eq, doTrace=0)
  boruta_results_eq[[metric]]  <- boruta.out
  # plot(boruta.out, cex.axis=.7, las=2, xlab="", main=metric) 
}
```

@@

```{r aggregate_features_5_eq}
features_top5 <- NULL
for(metric in names(boruta_results_eq)){
  features_top5 <- c(names(sort(apply(boruta_results_eq[[metric]]$ImpHistory, 2, median), decreasing = TRUE)[1:5]), features_top5)
}
knitr::kable(data.frame(top5 = sort(c(unique(features_top5)))))
```

Increasing to 10:

```{r aggregate_features_10_eq}
features_top10 <- NULL
for(metric in names(boruta_results_eq)){
  features_top10 <- c(names(sort(apply(boruta_results_eq[[metric]]$ImpHistory, 2, median), decreasing = TRUE)[1:10]), features_top10)
}
knitr::kable(data.frame(top10 = sort(c(unique(features_top10)))))
```

>>>>

```{r data_load, echo=FALSE}
load("~/ff-beta-release-matching/poc/matchIt/feature_selection.RData")
```

```{r}
filter_features <- function(df_train_sm){
  df_fs <- df_train_sm %>%
   select(-engagement) %>%
   select(-content_crashes) %>%
   select(-client_id) %>%
   select(-label_beta) %>%
   select(-label_release) %>%
   select(-is_release) %>%
   select(-app_version)
  return(df_fs)
}
```

@@

set.seed(1234)
df_train_sm <- df_train_encoder %>% sample_n(10000)
df_fs <- filter_features(df_train_sm)
```

@@

```{r apply_boruta}
boruta_results <- list()

for (metric in engagement){
  print(paste('Applying Boruta to ', metric))
  boruta.out <- Boruta(y = df_train_sm[[metric]], x=df_fs, doTrace=0)
  boruta_results[[metric]]  <- boruta.out
}
```

```{r boruta_plots, fig.height=7, fig.width=15}
for (metric in engagement){
  plot(boruta_results[[metric]], cex.axis=.7, las=2, xlab="", main=metric)
}
```

> Find the top 5 ranking features per metric, and add to a list.

```{r aggregate_features_5}
features_top5 <- NULL

for(metric in engagement){
  features_top5 <- c(names(sort(apply(boruta_results[[metric]]$ImpHistory, 2, median), decreasing = TRUE)[1:5]), features_top5)
}

top5 <- sort(c(unique(features_top5)))
x <- data.frame(top5)
kable(x) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

> Increasing to 10:

```{r aggregate_features_10}
features_top10 <- NULL

for(metric in engagement){
  features_top10 <- c(names(sort(apply(boruta_results[[metric]]$ImpHistory, 2, median), decreasing = TRUE)[1:10]), features_top10)
}

top10 <- sort(c(unique(features_top10)))
x <- data.frame(top10)
kable(x) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

@@

```{r var_def_equal}
df_beta <- df_train_encoder %>% 
  filter(label_beta == 1)
n_beta <- nrow(df_beta)
set.seed(1234)
df_rel <- df_train_encoder %>% 
  filter(label_beta == 0) %>%
  sample_n(n_beta)
set.seed(1234)
df_train_f_sm_eq <- df_rel %>%
  bind_rows(df_beta) %>%
   sample_n(10000)
df_fs_eq <- df_train_f_sm_eq %>%
   select(-engagement) %>%
   select(-content_crashes) %>%
   select(-client_id) %>%
   select(-label_beta) %>%
   select(-label_release) %>%
   select(-is_release) %>%
   select(-app_version)
```

```{r apply_boruta_equal}
boruta_results_eq <- list()
for (metric in engagement){
  print(paste('Applying Boruta to ', metric))
  boruta.out <- Boruta(y = df_train_f_sm_eq[[metric]], x=df_fs_eq, doTrace=0)
  boruta_results_eq[[metric]]  <- boruta.out
}
```

@@

> Find the top 5 ranking features per metric, and add to a list.

```{r aggregate_features_5_eq}
features_top5 <- NULL
for(metric in names(boruta_results_eq)){
  features_top5 <- c(names(sort(apply(boruta_results_eq[[metric]]$ImpHistory, 2, median), decreasing = TRUE)[1:5]), features_top5)
}

x <- data.frame(top5 = sort(c(unique(features_top5))))
kable(x) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```

> Increasing to 10:

```{r aggregate_features_10_eq}
features_top10 <- NULL
for(metric in names(boruta_results_eq)){
  features_top10 <- c(names(sort(apply(boruta_results_eq[[metric]]$ImpHistory, 2, median), decreasing = TRUE)[1:10]), features_top10)
}

x <- data.frame(top10 = sort(c(unique(features_top10))))
kable(x) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
scroll_box(width = "100%")
```
